---
title: "assigment5"
format: html
editor: visual
---

```{r}
library(ggplot2)
library(gganimate)
library(dplyr)
library(gapminder)
library(nycflights13)
```

# Dataset Context

## Gapminder Dataset

The Gapminder data set contains various information about other countries over periods of time inlcuding, GDP per capita. population, and life expectancy as well. The reason I chose this data set was because of the correlation between economic growth and health over periods of time.

## Flights Dataset

The flights dataset contains flight data from the year 2013 including months and delays. I chose this dataset to analyze how many delays tend to happen over time.

# Question 1

How has the relationship between per capita, GDP, and life expectancy changed with time ?

```{r}
p1 <- ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point() +
  scale_x_log10() +
  labs(title = "Year: {frame_time}") +
  transition_time(year)

animate(p1)
```

# Explanation 1

Countries with a higher overall GDP tend to have higher life expectancies. With time most of the countires will improve in both areas.

# Question 2

How many flight delays increase or decrease across the months?

```{r}
flights_summary <- flights %>%
  group_by(month) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE))

p2 <- ggplot(flights_summary, aes(month, delay)) +
  geom_col() +
  labs(title = "Month: {frame_time}") +
  transition_states(month)

animate(p2)
```
